\chapter{Introduction}
\label{chap:introduction}

\section{Background}

In today's digital age, fashion trends are heavily influenced by celebrities, influencers, and social media figures. Images of individuals showcasing unique and appealing outfits frequently circulate on social platforms, movies, and public appearances. These images often inspire personal fashion choices among users. However, translating such visual inspiration into actual purchases remains a significant challenge.

The conventional process of identifying and purchasing similar fashion items involves manual searching across various e-commerce platforms, which can be time-consuming, inaccurate, and frustrating. Despite the rapid growth of online shopping, there is still a noticeable gap between user inspiration and product discovery in the fashion retail sector.

To address this, the integration of advanced technologies like object detection and visual similarity search into e-commerce platforms has emerged as a promising solution. Such technologies can bridge the gap between inspiration and purchase, creating a more intuitive and efficient shopping experience.

\section{Motivation}

The fashion retail industry is constantly seeking innovative ways to enhance user experience and increase sales conversions. Enabling users to effortlessly find products inspired by images they admire presents a transformative opportunity.

Imagine a user uploading a photo of a celebrity or an influencer wearing a desirable outfit, and the system automatically detecting the fashion items and retrieving visually similar products from the store's inventory. This would eliminate the manual, often tedious, search process and make fashion shopping more accessible, personalized, and enjoyable.

With the proliferation of mobile devices, camera technologies, and improvements in machine learning algorithms, the time is ripe to develop an AI-based shopping assistant tailored specifically for fashion products. Our project is driven by the goal of empowering users to bridge their fashion inspirations with actual products through advanced image and text-based search technologies.

\section{Objectives}

The primary objectives of the \textbf{AI Shopping Assistant} project are as follows:

\begin{itemize}
    \item To develop an application that assists users in finding their desired fashion items easily and efficiently.
    \item To implement an AI powered text-based product search system that enables users to search for fashion items using natural language like ("show me some black and green running shoes").
    \item To design and integrate an image-based product search functionality that allows users to upload an image and retrieve visually similar fashion products from the inventory.
    \item To leverage object detection and visual similarity algorithms specifically tailored to the domain of fashion products.
\end{itemize}

By achieving these objectives, the project aims to revolutionize the online fashion shopping experience, making it more aligned with modern user expectations and technological advancements.


